Sketch Recognition using Particle Swarm Optimization Algorithm

### 2009-04-12 trunk/draft/chapterSurvey.tex
after correctly splled now look at any thing else

\chapter{Literature Survey}
\label{sec:survey}
% this will contians only the finished work .... and my comments to view the old survey see the file
%draft_chaptersurvey 
%\section{Introduction }
%\label{sec:Introduction}
The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams. Current technology and the increae in computuing power have increased the abilities of such interfaces and allowed the implementaion of fully sketching systems. Over the years, many other sketching based system have been developed, elthough throught this history most system foucused on building the interactive interfaces to assist users while entering information to the computer. To demonstrate our system inspiration we will first present current research in the area of sketch recognition. Fristly, preprosessing and segmentaiton methods will be discussed in details. Later, symbol recognition methods will be presented and analysied in Section \ref{sec:symbolrecogntion}.

% i need to write a history then small introudction to the survey...
%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
\section{Preprocessing and Segmentation}
\label{sec:preprocessing}
%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
 % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
  In a system that supports free hand sketching, sketching should be as natural as possible. This means that the user can draw symbols using single or multiple strokes without constrains. The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper. But sketchs are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users. This led researchers to perform sort of preprocessing before starting the sketch recognition task itself. The next sections explain different preprocessing techniques.

@@

 Agar et al. \cite{polygonfeedback31} used the timestamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maxima than the speed information. He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data.

@@

\section{Symbol Recogntion}
\label{sec:symbolrecogntion}

@@

\subsubsection{Guesture Recognition}
\label{sec:GuestureRecognition}

@@

 Systems that employ such representation must first preprocess strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods). Recognition can then be treated as sub graph matching problem.

@@

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
	\caption{square scemantic graph}
	\label{fig:squarescematic}
\end{figure}

@@

 \paragraph{Visual Describtive Languages:}
 \label{sec:VisualDescribtiveLanguages}

 Avoiding graph matching problems, research tried to use languages that discribe shapes instead of graphs. Visual describtive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives. Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}. Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.

 Several visual grammer were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}. \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified. The system  automatically generates java code for the recognizer for each object. The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.  These systems are less computaitonal expensive than graph matching approach but they also operate based on the output of the segmentation process. Any error in the segmentation process lead to huge error while parsing the whole sketch. Moreover, symbols cannot share strokes between them which infleunce and restrict how the user draw symbols. 
 % complexity of the graph matching problem 
\begin{figure}
	\centering
		\subfigure[An Arrow]
			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
			\hfill
		\subfigure[Arrow grammar definition]
		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
	\caption[Arrow definition]{}
		\label{fig:arrowladderdef1}
\end{figure}

@@

 Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the preimeter to the area.  These global feature can be used to distinguish between classes. For example, a rectangle can be distinguish from a triangle by looking to the raito of area of bounding box to the area of the convex hull. The triangle ratio will be much lower than that of the rectangle.

@@

%Zernike moments were used by Hse et al. in \cite{ToADD}. 
Hse et al. \cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property. Zernike moemnts are orthognals moments which define the distribution of point in the input space. The higher the order of the moment the finer the detail level the point distribution represents. Hse et al. \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments. Early evluation of Zernike moemnts showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. In our system we used Zernike moemnts as part of our feature vector. A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.

@@

 Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have. These properties dose not depend on the order or number of strokes and thus are only based on apperance of the shape not how the user draw it. However, individual details of the shape are impossible to distinguish using these properties. For example, it is impoissible to use  to  differentiate between a current source symbole (circle contiaing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).  Finear details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distiguished using this method. These symbols requires a system to represent different levels of shape details.

@@

In the third approach, the recognition foucus on appearance of shapes contrary to individual strokes or global shape properties. Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the  prototypes in the database. Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system. Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes. He used two measures (Hausdor distance)  to measure similarity and two coeficients (Yule, Tanimoto) to measure dissimilarity. His approach allows the system to recongized a shape using only a single training sample as it is based on the shape prototype. However, the shape protoytpe must represent all  transformation and variation the input shape can have. Moreover, they restrict the image representation to 48x48 pixel due to computationa reasons. This restriction leads to elimintating some of the fine details that may be needed to distinguish some shape classes. There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.

@@

In all of these system only \cite{SketchRead2007} using stroke based recognitiion have been implemented to the fully creat a system that free draw sketch and understand them. This is because the emphasis on the interactive side of the sketch sytem which result in large difficultes in processing free hand sketches. Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be groubed. These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}. This let them to temporally group strokes and then recognize them based on their strokes or apperance properties. \cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketchs.

Recogntion based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}. The principle problem in these systems is the computatnal complexity of the matching between shape description and geometeric primitves produced in the segmentaiton process. Natural sketch are dificult to segment and group to create reliable fragment of geometeric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, overtraced strokes and sharing of strokes between different symbols.  Since these systems are depenable on the segmenation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.

To avoid most problems in such system, our system uses hybird of both global shape and hierarchical discribtion methods. Currenlty, our system recognize on sysmbols at a time. Each symbol is segmented and hierarchical features are computed.  Features  that are represent the global discription of the shape are also generated and appended to the hierarchical featusres. The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.

\chapter{Particle Swram Optimization}
\label{sec:ParticleSwramOptimization}

\section{Swram Optimization Review}
\label{sec:ParticleSwramOptimizationReview}

@@

Modeling how a biological swarm has been recentlsy studied by biologist and computer scientists. Scientist in the field or artifical life need to understand how such "social anaimals" interact, achieve goals and evolove. Moreover, swarm behaviours that achieves some kind of swram intelligance is increasly intersting for engineers as it they can be applied in optimization. Looking from a higher view, agents in a swarm are cooperating to achieve some proposeful behaviour and achieve some common goal. For example, ant groups looking for food source. Large gourps of simple independent agents seems to generate this behaviour in what is called 'collective intelligence'. Each agent has a set of simple rules to follow, and accourding to the interaction of the entrie group the objective is met\cite{swarmGeneral}. %with other neighbourhood agents.

Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds. Each boid is a simple agent that navigate according to his perception and a set of basic rules.  Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behvaiour and used in cooperative transportation. Ants leaves a trail of its movement, so when an ant find food it cannot cary alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.   Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to corectly models ants behavour with a methods of task modeling.

Image processing is one of the displines that used Swarm Intelligance. Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}. Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm alogrithm (PSO) the algorithm had been used in various image processing sytem as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.  PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentaiton process.

>>>>

\chapter{Literature Survey}
\label{sec:survey}
% this will contians only the finished work .... and my comments to view the old survey see the file
%draft_chaptersurvey 
%\section{Introduction }
%\label{sec:Introduction}
The history of pen and sketch based computing goes back to 1963 with Ivan Sutherland's SKETCHPAD \cite{sutherlandsketch}, which used a light pen to draw on the monitor to create circuit diagrams. Current technology and the increase in computing power have increased the abilities of such interfaces and allowed the implementation of fully sketching systems. Over the years, many other sketching based system have been developed, although through this history most system focused on building the interactive interfaces to assist users while entering information to the computer. To demonstrate our system inspiration we will first present current research in the area of sketch recognition. Firstly, preprocessing and segmentation methods will be discussed in details. Later, symbol recognition methods will be presented and analyzed in Section \ref{sec:symbolrecogntion}.

% i need to write a history then small introudction to the survey...
%  i have to state that the next  sections will generally group segmentation and symbol recognition task. 
\section{Preprocessing and Segmentation}
\label{sec:preprocessing}
%this section should describe how others did the first preprocessing of the problem. i have to summarize what i wrote before. 
 % first paragrah about how stroke is a messing thing and naturally it is ambigues. 
  In a system that supports free hand sketching, sketching should be as natural as possible. This means that the user can draw symbols using single or multiple strokes without constrains. The goal of a sketch system is making the user feel that using the using the system is nearly the same as using pencil and paper. But sketches are incomplete drawings that are sloppy and messy, for example figures \ref{fig:Overshootandundershoot} and \ref{fig:overstroked} show examples of messy sketches drawn by users. This led researchers to perform sort of preprocessing before starting the sketch recognition task itself. The next sections explain different preprocessing techniques.

@@

 Agar et al. \cite{polygonfeedback31} used the time stamp for each point and compute time difference between every two successive point to help him finding corners as it provides more distinct maximum than the speed information. He mentions that the pointing device (for example the mouse) sampling rate is the reason for this phenomena. The points are sampled at regular time intervals while the pointing device is moving. There are no points while the pointing device is stationary. This leads to a nearly constant time difference between samples while the pen is moving and large difference while the pen is stationary. Contrary to speed information, where the users draw with variant speeds the speed information has a lot of noise in the data.

@@

\section{Symbol Recognition}
\label{sec:symbolrecogntion}

@@

\subsubsection{Gesture Recognition}
\label{sec:GuestureRecognition}

@@

 Systems that employ such representation must first pre-process strokes to break them into geometric primitives (see section \ref{sec:preprocessing} for a review on preprocessing methods). Recognition can then be treated as sub graph matching problem.

@@

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{../../neededfiles/Figures/squarescematic.eps}
	\caption{square semantic graph}
	\label{fig:squarescematic}
\end{figure}

@@

 \paragraph{Visual Descriptive Languages:}
 \label{sec:VisualDescribtiveLanguages}

 Avoiding graph matching problems, research tried to use languages that describe shapes instead of graphs. Visual descriptive languages represents the shape into tokens of geometric primitives or relations between smaller  geometric primitives. Figures \ref{fig:arrow1} and \ref{fig:arr2} show a regular arrow with 2 shafts and head and its definition in Ladder language \cite{Ladder30}. Shape recognition is then a question of parsing the sketch document into a recognizer that parse strokes according to the constructed grammar.

 Several visual grammar were introduced \cite{statisticalparsing26,Ladder30,GenericHMM28}. \cite{Ladder30,GenericHMM28}  Used visual description language which will generate the information of the structure and constrain of each object classified. The system  automatically generates Java code for the recognizer for each object. The generated code cycles through the strokes drawn to check if any of the recognizers generated had interpreted any symbols partially or completely.  These systems are less computational expensive than graph matching approach but they also operate based on the output of the segmentation process. Any error in the segmentation process lead to huge error while parsing the whole sketch. Moreover, symbols cannot share strokes between them which influence and restrict how the user draw symbols. 
 % complexity of the graph matching problem 
\begin{figure}
	\centering
		\subfigure[An Arrow]
			{\label{fig:arrow1}\includegraphics[scale=0.75]{../../neededfiles/Figures/arrowladderdef1.eps}}
			\hfill
		\subfigure[Arrow grammar definition]
		 {\label{fig:arr2}\includegraphics[scale=0.88]{../../neededfiles/Figures/arrowladderdef2.eps}}
	\caption[Arrow definition]{}
		\label{fig:arrowladderdef1}
\end{figure}

@@

 Several systems \cite{DiagramOfflineConvexHull,MulitStrokeConvexHull} presented features like ratio of bounding box to convex hull area and the ratio of the perimeter to the area.  These global feature can be used to distinguish between classes. For example, a rectangle can be distinguish from a triangle by looking to the ratio of area of bounding box to the area of the convex hull. The triangle ratio will be much lower than that of the rectangle.

@@

%Zernike moments were used by Hse et al. in \cite{ToADD}. 
Hse et al. \cite{HeloiseBeautification,zernike61} proposed using the magnitudes of Zernike moments as a shape property. Zernike moments are orthogonal moments which define the distribution of point in the input space. The higher the order of the moment the finer the detail level the point distribution represents. Hse et al. \cite{HeloiseBeautification,zernike61} used a SVM classifier to train the features vector calculated from the moments. Early evaluation of Zernike moments showed that they are invariant to both reflection and rotation of the input shape \cite{ZerMomentOrthogonal}. In our system we used Zernike moments as part of our feature vector. A detailed comparison between Zernike moments and other features are demonstrated in Section \ref{sec:featuresComparisions}.

@@

 Systems based on Global shape properties distinguish shapes by learning the different properties each shape class tend to have. These properties dose not depend on the order or number of strokes and thus are only based on appearance of the shape not how the user draw it. However, individual details of the shape are impossible to distinguish using these properties. For example, it is impossible to use  to  differentiate between a current source symbol (circle containing and arrow Figure \ref{fig:CurrentSource}) and a Ac Source (circle containing a tilate Figure \ref{fig:acSource}).  Finer details of the symbols are not represented in this approach, thus shapes that have smalls details or only significant conceptual variation are not distinguished using this method. These symbols requires a system to represent different levels of shape details.

@@

In the third approach, the recognition focus on appearance of shapes contrary to individual strokes or global shape properties. Kara \cite{imagetrainable48} used this approach while presenting a system that matches the input shape into a one of the  prototypes in the database. Figure \ref{fig:template} shows an example of prototypes of some of the shapes in the system. Input shapes are first pose and scale normalized then four image similarity measure are used to match input shapes to the database prototypes. He used two measures (Hausdor distance)  to measure similarity and two coefficients (Yule, Tanimoto) to measure dissimilarity. His approach allows the system to recognized a shape using only a single training sample as it is based on the shape prototype. However, the shape prototype must represent all  transformation and variation the input shape can have. Moreover, they restrict the image representation to 48x48 pixel due to computational reasons. This restriction leads to eliminating some of the fine details that may be needed to distinguish some shape classes. There results on isolated symbols seems promising but the system have not beet tested on full sketch that tends to have high overlapping between symbols.

@@

In all of these system only \cite{SketchRead2007} using stroke based recognition have been implemented to the fully creat a system that free draw sketch and understand them. This is because the emphasis on the interactive side of the sketch sytem which result in large difficulties in processing free hand sketches. Methods based on gesture and global stroke properties produce systems that are dependent on assumptions about how strokes can be grouped. These systems assume that users will finish drawing a shape before moving to the next one \cite{multistrokeEvaluation,Cali63,geometrydomain49}. This let them to temporally group strokes and then recognize them based on their strokes or appearance properties. \cite{AlvaradoDigital} argued that this assumption dose not hold in natural sketches.

Recognition based on strokes using hierarchical shape description is applied in \cite{SketchRead2007}. The principle problem in these systems is the computational complexity of the matching between shape description and geometric primitives produced in the segmentation process. Natural sketch are difficult to segment and group to create reliable fragment of geometeric primitives. This is partially due to the increase of noise and other phenomena such as overlapped, over traced strokes and sharing of strokes between different symbols.  Since these systems are dependable on the segmentation results, they either performe recognition using undependable segmentation results or consider different ways to segment the input sketch.

To avoid most problems in such system, our system uses hybrid of both global shape and hierarchical description methods. Currently, our system recognize on symbols at a time. Each symbol is segmented and hierarchical features are computed.  Features  that are represent the global description of the shape are also generated and appended to the hierarchical features. The recognition is done using SVM Linear classifier as it proved to be one of the best learning algorithms.

\chapter{Particle Swarm Optimization}
\label{sec:ParticleSwramOptimization}

\section{Swarm Optimization Review}
\label{sec:ParticleSwramOptimizationReview}

@@

Modeling how a biological swarm has been recently studied by biologist and computer scientists. Scientist in the field or artificial life need to understand how such "social animals" interact, achieve goals and evolve. Moreover, swarm behaviors that achieves some kind of swarm intelligence is increasingly interesting for engineers as it they can be applied in optimization. Looking from a higher view, agents in a swarm are cooperating to achieve some purposeful behavior and achieve some common goal. For example, ant groups looking for food source. Large groups of simple independent agents seems to generate this behavior in what is called 'collective intelligence'. Each agent has a set of simple rules to follow, and according to the interaction of the entire group the objective is met\cite{swarmGeneral}. %with other neighbourhood agents.

Reynolds \cite{swarmflock} created 'boid' which simulated the motion of a flock of birds. Each boid is a simple agent that navigate according to his perception and a set of basic rules.  Kube and Zhange \cite{antsimulation} introduced a simulation model for ants behavior and used in cooperative transportation. Ants leaves a trail of its movement, so when an ant find food it cannot carry alone, it can easily trace back to the colony and then a group of ants can track the trail and bring the food back.   Kube and Zhange \cite{antsimulation,swarmGeneral} system seems to correctly models ants behavior with a methods of task modeling.

Image processing is one of the disciplines that used Swarm Intelligence. Ant colony was widely used to segment objects in \cite{antSegmentation,antthreshold,SwarmMedical}. Since 1995 when Kennedy and Eberhart \cite{PSOFirst,PSO2} introduced particle swarm algorithm (PSO) the algorithm had been used in various image processing system as an optimization  algorithm \cite{SwarmImageProcessing,Swarmimageprocessing2,Colorquantization,SwarmMedical}.  PSO algorithm was used in \cite{ToADD} to classify pixel to simplify the object segmentation process.
